DDS4Chl: A dataset of drought indices of water demand and supply and its impact on vegetation for continental Chile

### 2024-01-11 article/paper_drought_Chile2023.qmd
agregar mapa koppe-geiger al área de estudio

In this research, we aim to analyze the impact of drought on different types of land cover classes in continental Chile by combining environmental variables such as biomass productivity, and water demand and supply gathered from earth observation products. The specific objectives of the study are: i) to analize the trend on multi-scalar drought indices for water demand/supply, soil moisture and vegetation productivity for 1981–2023/2001-2023; ii) to evaluate LULC change for 2001–2021 and its relation to drought indices; iii) to analyze the relationship of a proxy of biomass (zcNDVI) with drought indices; and iv) to assess if the observed changes in the drought indices are linked to TWS.

@@

## Drought indices

##

##	Drought indices for water demand/supply, soil moisture and vegetation productivity

>>>>

In this research, we present the dataset DDS4Chl, which provides climate variables and drought indices of water demand and  supply and vegetation productivity for continental Chile since 1981 at a monthly frequency. Those were gathered from the earth observation products ERA5-Land and MODIS. Then, we used DDS4Chl to analyze the impact of drought on different types of land cover classes in continental Chile. The specific objectives of the study are: i) to analyze the trend on multi-scalar drought indices for water demand and supply, soil moisture, and vegetation productivity for 1981–2023/2001–2023; ii) to evaluate LULC change for 2001–2021 and its relation to drought indices; iii) to analyze the relationship of a proxy of vegetation productivity (zcNDVI) with drought indices of water demand and supply and soil moisture; and iv) to assess if the observed changes in the drought indices are linked to TWS.

@@

##	Drought indices for water demand and supply, soil moisture and vegetation productivity

### 2024-01-11 article/paper_drought_Chile2023.tex
agregar mapa koppe-geiger al área de estudio

\begin{keyword}
    drought \sep land cover change \sep 
    satellite
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, enhanced, boxrule=0pt, interior hidden, breakable]}{\end{tcolorbox}}\fi

@@

To evaluate meteorological drought (i.e., water supply), the World
Meteorological Organization (WMO; \citep{WMO2012}) recommends the
Standardized Precipitation Index (SPI; \citep{Mckee1993}), a multiscalar
drought index that allows to monitor precipitation deficits from short-
to long-term. Following the same approach, \citep{Vicente-Serrano2010}
incorporates into the SPI the effect of temperature through the use of
potential evapotranspiration, thus proposing the SPEI (Standardized
Precipitation Evapotranspiration Index). Similarly, to evaluate solely
the evaporative demand driven by temperature, \citep{Hobbins2016} and
\citep{McEvoy2016} came up with the Evaporative Demand Drought Index
(EDDI). For vegetation, in a similar manner as the SPI, SPEI and EDDI;
~\citep{Zambrano2018} proposed the zcNDVI, a standardized anomaly of the
cumulative Normalized Difference Vegetation Index (NDVI), which could be
acumulated over the growing season or any period (e.g., months),
resulting in a multiscalar drought index. For soil moisture, several
drought indices exist, such as the Soil Moisture Deficit Index (SDMI) a
normalized index \citep{Narasimhan2005} and the Soil Moisture
Agricultural Drought Index (SMADI) \citep{Souza2021} which is a
normalized index using vegetation, land surface temperature, and a
vegetation condition index (VCI, \citep{Kogan1995}). From TWS, we can
estimate the standardized terrestrial water storage index (STI)
\citep{Cui2021}, a standardized anomaly that follows the methodology of
the SPI, SPEI, EDDI, and zcNDVI. Thereby, we have drought indices for
water supply, demand, and storage, which can help to make a
comprehensive assessment of drought.

In this research, we aim to analyze the impact of drought on different
types of land cover classes in continental Chile by combining
environmental variables such as biomass productivity, and water demand
and supply gathered from earth observation products. The specific
objectives of the study are: i) to analize the trend on multi-scalar
drought indices for water demand/supply, soil moisture and vegetation
productivity for 1981--2023/2001-2023; ii) to evaluate LULC change for
2001--2021 and its relation to drought indices; iii) to analyze the
relationship of a proxy of biomass (zcNDVI) with drought indices; and
iv) to assess if the observed changes in the drought indices are linked
to TWS.

@@

\hypertarget{drought-indices}{%
\subsection{Drought indices}\label{drought-indices}}

\hypertarget{section}{%
\subsection{}\label{section}}

\hypertarget{drought-indices-for-water-demandsupply-soil-moisture-and-vegetation-productivity}{%
\subsection{Drought indices for water demand/supply, soil moisture and
vegetation
productivity}\label{drought-indices-for-water-demandsupply-soil-moisture-and-vegetation-productivity}}

>>>>

\begin{keyword}
    drought \sep land cover change \sep 
    satellite
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, interior hidden, enhanced, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, breakable, sharp corners]}{\end{tcolorbox}}\fi

@@

To evaluate meteorological drought (i.e., water supply), the World
Meteorological Organization (WMO; \citep{WMO2012}) recommends the
Standardized Precipitation Index (SPI; \citep{Mckee1993}), a multiscalar
drought index that allows to monitor precipitation deficits from short-
to long-term. Following the same approach, \citep{Vicente-Serrano2010}
incorporates into the SPI the effect of temperature through the use of
potential evapotranspiration, thus proposing the SPEI (Standardized
Precipitation Evapotranspiration Index). Similarly, to evaluate solely
the evaporative demand driven by temperature, \citep{Hobbins2016} and
\citep{McEvoy2016} came up with the Evaporative Demand Drought Index
(EDDI). For vegetation, in a similar manner as the SPI, SPEI and EDDI;
\citep{Zambrano2018} proposed the zcNDVI, a standardized anomaly of the
cumulative Normalized Difference Vegetation Index (NDVI), which could be
acumulated over the growing season or any period (e.g., months),
resulting in a multiscalar drought index. For soil moisture, several
drought indices exist, such as the Soil Moisture Deficit Index (SDMI) a
normalized index \citep{Narasimhan2005} and the Soil Moisture
Agricultural Drought Index (SMADI) \citep{Souza2021} which is a
normalized index using vegetation, land surface temperature, and a
vegetation condition index (VCI, \citep{Kogan1995}). From TWS, we can
estimate the standardized terrestrial water storage index (STI)
\citep{Cui2021}, a standardized anomaly that follows the methodology of
the SPI, SPEI, EDDI, and zcNDVI. Thereby, we have drought indices for
water supply, demand, and storage, which can help to make a
comprehensive assessment of drought.

In this research, we present the dataset DDS4Chl, which provides climate
variables and drought indices of water demand and~ supply and vegetation
productivity for continental Chile since 1981 at a monthly frequency.
Those were gathered from the earth observation products ERA5-Land and
MODIS. Then, we used DDS4Chl to analyze the impact of drought on
different types of land cover classes in continental Chile. The specific
objectives of the study are: i) to analyze the trend on multi-scalar
drought indices for water demand and supply, soil moisture, and
vegetation productivity for 1981--2023/2001--2023; ii) to evaluate LULC
change for 2001--2021 and its relation to drought indices; iii) to
analyze the relationship of a proxy of vegetation productivity (zcNDVI)
with drought indices of water demand and supply and soil moisture; and
iv) to assess if the observed changes in the drought indices are linked
to TWS.

@@

\hypertarget{drought-indices-for-water-demand-and-supply-soil-moisture-and-vegetation-productivity}{%
\subsection{Drought indices for water demand and supply, soil moisture
and vegetation
productivity}\label{drought-indices-for-water-demand-and-supply-soil-moisture-and-vegetation-productivity}}

### 2024-01-11 conference/IGARSS2024.qmd
agregar mapa koppe-geiger al área de estudio

To evaluate meteorological drought (i.e., water supply), the World Meteorological Organization (WMO; @WMO2012) recommends the Standardized Precipitation Index (SPI; @McKee1993), a multiscalar drought index that allows to monitor precipitation deficits from short- to long-term. Following the same approach, @Vicente-Serrano2010 incorporates into the SPI the effect of temperature through the use of potential evapotranspiration, thus proposing the SPEI (Standardized Precipitation Evapotranspiration Index). Similarly, to evaluate solely the evaporative demand driven by temperature, @Hobbins2016 and @McEvoy2016 came up with the Evaporative Demand Drought Index (EDDI). To assess the impact of water supply and demand on vegetation, Zambrano (2018) proposed the zcNDVI. This is a standardized anomaly of the cumulative Normalized Difference Vegetation Index (NDVI). It's similar to the SPI, SPEI, and EDDI. The zcNDVI could be accumulated over the growing season or any period (e.g., months). It results in a multiscalar drought index. Several drought indices exist for soil moisture. These include the Soil Moisture Deficit Index (SDMI), a normalized index [@Narasimhan2005], and the Soil Moisture Agricultural Drought Index (SMADI) [@Souza2021]. SMADI is a normalized index that uses vegetation, land surface temperature, and a vegetation condition index (VCI, [@Kogan1995]).

>>>>

To evaluate meteorological drought (i.e., water supply), the World Meteorological Organization (WMO; @WMO2012;@Hayes2011) recommends the Standardized Precipitation Index (SPI; @McKee1993), a multiscalar drought index that allows to monitor precipitation deficits from short- to long-term. Following the same approach, @Vicente-Serrano2010 incorporates into the SPI the effect of temperature through the use of potential evapotranspiration, thus proposing the SPEI (Standardized Precipitation Evapotranspiration Index). Similarly, to evaluate solely the evaporative demand driven by temperature, @Hobbins2016 and @McEvoy2016 came up with the Evaporative Demand Drought Index (EDDI). To assess the impact of water supply and demand on vegetation, Zambrano (2018) proposed the zcNDVI. This is a standardized anomaly of the cumulative Normalized Difference Vegetation Index (NDVI). It's similar to the SPI, SPEI, and EDDI. The zcNDVI could be accumulated over the growing season or any period (e.g., months). It results in a multiscalar drought index. Several drought indices exist for soil moisture. These include the Soil Moisture Deficit Index (SDMI), a normalized index [@Narasimhan2005], and the Soil Moisture Agricultural Drought Index (SMADI) [@Souza2021]. SMADI is a normalized index that uses vegetation, land surface temperature, and a vegetation condition index (VCI, [@Kogan1995]).
